---
title: "Abisko_Chemistry_Merge"
author: "Matilda Andersson"
date: "2026-06-24"
output: html_document
---

load packages
```{r}
library(tidyverse)
library(readxl)
```

load data - this now just uploads the entire Abisko folder with all the data
IMPORTANT - it's important that the file names follow the naming conventions so that everything uploaded makes sense
```{r}
list.files(path = "ProcessedData/Abisko", 
           pattern = "\\.csv$", 
           full.names = TRUE) %>%
  set_names(~ basename(.) %>% str_remove("\\.csv$")) %>%
  map(read_csv) %>%
  list2env(envir = .GlobalEnv)

rm(TagTag_ANS_ALM_2022) #this is for Almbergasjön which is no longer being updated
```

```{r}
# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "ANS" #STATION Abbreviation here
LakeSamplingLocation <- "LHL" #sampling code here
EndYear <- 2025 #END YEAR HERE
```

YOU STOPED HERE -- need the TagTag for Långa Harrsjön before you continue
convert LOD to 99999 and NaN to NA and convert everything to numeric 
```{r}
TagTag <- TagTag %>%
  mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.numeric))

df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
df_names <- df_names[df_names != "TagTag"]

for(df_name in df_names) {
  assign(df_name, get(df_name) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.character)) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(. == "LOD", "99999", .))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ suppressWarnings(as.numeric(.)))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(is.nan(.), NA_real_, .)))
  )
}
```

merge any overlapping data sets
```{r}
#NP
NP_all <- bind_rows(mget(ls(pattern = "^NP"))) |> 
  distinct()
rm(list = setdiff(ls(pattern = "^NP"), "NP_all"))

#DOC
DOC_TOTN_all <- bind_rows(mget(ls(pattern = "^DOC")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^DOC"), "DOC_TOTN_all"))

#Anions
Anions_all <- bind_rows(mget(ls(pattern = "^Anions")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Anions"), "Anions_all"))

#Cations
Cations_all <- bind_rows(mget(ls(pattern = "^Cations")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Cations"), "Cations_all"))

```


merge data - keep everything from TagTag leave NAs in the other locations
```{r}
TagTag <- TagTag %>%
  left_join(DOC_TOTN_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(NP_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Anions_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Cations_all, by = "SampleID")

rm(DOC_TOTN_all, NP_all, Anions_all, Cations_all)

#just for abisko - drop F until things are re-run - there was a problem with a difference in magnitude between the two labs
# TagTag <- TagTag %>% 
#   select(-"F")

rm(EC_pH_Asa_2022_2025)

```
