---
title: "Anions"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

install.packages("usethis")  # if needed
usethis::use_gitignore()
usethis::use_git()

load packages
```{r}
library(tidyverse)
library(readxl)
```

#Uppsala results
You need to use code specifically for UU because they format their data in a specific way and have specific units
CHECK SOMETHING - check in the header that everything is mg/L -- otherwise you need a different conversion 
CHECK that they have decimals as . and not ,
CHECK that they don't have any duplicates (labeled duplicate in the )

load data - sometimes the filenames are different so you cant fully automate
*CHANGE SOMETHING MANUALLY HERE*
```{r}
rm(list = ls())

rawfile <- "anions_UU_Skogaryd_2025" #ENTER YOUR FILENAME HERE

filepath <- paste0("LabResults/Anions/", rawfile, ".xlsx")
df <- read_excel(filepath, sheet = "RESULTS")

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "Skogaryd" #STATION NAME HERE
StartYear <- 2025 #START YEAR HERE
EndYear <- 2025 #END YEAR HERE
```

**check results** - if something other than mg/L shows up you will need to do a conversion
```{r}
unique(as.character(unlist(df[1, ])))
```

drop unnecessary columns and rows
```{r}
#drop anything below the data
df <- df %>%
  filter(row_number() <= max(which(complete.cases(.))))

#drop the column SO4 (you only need SO4-S) --- you only need this if you haven't manually removed this column when fixing the excel sheet
df <- df %>%
  select(-any_of("SO4"))

# the column names are correct and are what is used on the portal so we will just drop row 2
df <- df %>%
  slice(-1)

#change the column name of the first column
df <- df %>%
  rename(SampleID = 1)

df <- df %>%
  mutate(SampleID = sub(".*?\\s+", "", SampleID))
```

remove values below LOD 
We are going to change anything with a < to to 0.000 and then change that to LOD after doing the unit conversion
```{r}
df <- df %>%
  mutate(across(everything(), ~ ifelse(grepl("^<", .), 0.000, .)))
```

```{r}
df <- df %>%
  # convert all columns except the first to numeric
  mutate(across(-1, as.numeric))
```

for the Uppsala data you need to convert everything from mg/L to ug/L
you should double check the original file that it's all mg
```{r}
#then recalculate
df <- df %>%
  mutate(across(-SampleID, ~ as.numeric(.) * 1000))
```

add LOD
```{r}
df <- df %>%
  mutate(across(-SampleID, ~ ifelse(. == 0.000, "LOD", .)))
```


Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/Anions_", StationName, "_", StartYear,"_",EndYear, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE, quote = FALSE)
```
