---
title: "Asa Chemistry"
author: "Matilda Andersson"
date: "2026-04-21"
output: html_document
---

Station specific code to deal with the unique formatting at the station level and the removal of any outliers specific to the station

The way that this is written you are just adding the new files when you load data so for each new upload it's generating a whole new file using all the past data. The thing that you will notice is that in the figures there are many overlaping points with both data from the portal and new data plotted on top of each other (should be the same point). Only data that has been loaded/added for the first time doesn't have overlapping points (and the really old data that I don't have the raw data for). This shouldn't be a problem. 

load packages
```{r}
library(tidyverse)
library(readxl)
```

#load data - 
add to this when new data becomes avaliable 
```{r}
TagTag <- read_excel("TagTag/ASA/TagTag_Asa.xlsx")

#Anions
Anions_Asa_2020_2021 <- read_csv("ProcessedData/Anions_Asa_2020_2021.csv")
Anions_Asa_2021_2022 <- read_csv("ProcessedData/Anions_Asa_2021_2022.csv")
Anions_Asa_2022_2023 <- read_csv("ProcessedData/Anions_Asa_2022_2023.csv")
Anions_Asa_2024_2024 <- read_csv("ProcessedData/Anions_Asa_2024_2024.csv")
Anions_Asa_2024_2025 <- read_csv("ProcessedData/Anions_Asa_2024_2025.csv")

#Cations
Cations_Asa_2017_2022 <- read_csv("ProcessedData/Cations_Asa_2017_2022.csv")
Cations_Asa_2022_2023 <- read_csv("ProcessedData/Cations_Asa_2022_2023.csv")
Cations_Asa_2023_2024 <- read_csv("ProcessedData/Cations_Asa_2023_2024.csv")
Cations_Asa_2024_2025 <- read_csv("ProcessedData/Cations_Asa_2024_2025.csv")

#DOC and TOTN
DOC_TOTN_Asa_2022_2022 <- read_csv("ProcessedData/DOC_TOTN_Asa_2022_2022.csv")
DOC_TOTN_Asa_2023_2023 <- read_csv("ProcessedData/DOC_TOTN_Asa_2023_2023.csv")
DOC_TOTN_Asa_2024_2024 <- read_csv("ProcessedData/DOC_TOTN_Asa_2024_2024.csv")
DOC_TOTN_Asa_2025_2026 <- read_csv("ProcessedData/DOC_TOTN_Asa_2025_2026.csv")

#Nitrogen and Phosphorous NP
NP_Asa_2022_2022 <- read_csv("ProcessedData/NP_Asa_2022_2022.csv")
NP_Asa_2023_2023 <- read_csv("ProcessedData/NP_Asa_2023_2023.csv")
NP_Asa_2023_2024 <- read_csv("ProcessedData/NP_Asa_2023_2024.csv")
NP_Asa_2024_2025 <- read_csv("ProcessedData/NP_Asa_2024_2025.csv")
NP_Asa_2025_2025 <- read_csv("ProcessedData/NP_Asa_2025_2025.csv")

#EC and pH
EC_pH_Asa_2022_2025 <- read_csv("ProcessedData/EC_pH_Asa_2022_2025.csv")

```

```{r}
# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "ASA" #STATION Abbreviation here
LakeSamplingLocation <- "FER" #sampling code here
EndYear <- 2025 #END YEAR HERE
```

convert LOD to 99999 and NaN to NA and convert everything to numeric 
```{r}
TagTag <- TagTag %>%
  mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.numeric))

df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
df_names <- df_names[df_names != "TagTag"]

for(df_name in df_names) {
  assign(df_name, get(df_name) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.character)) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(. == "LOD", "99999", .))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ suppressWarnings(as.numeric(.)))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(is.nan(.), NA_real_, .)))
  )
}
```

#merge any overlapping data sets
only distinct rows - if there is a copy paste error when the data is being compiled you could have two duplicate rows with the same data for the same date
```{r}
#NP
NP_all <- bind_rows(mget(ls(pattern = "^NP"))) |> 
  distinct()
rm(list = setdiff(ls(pattern = "^NP"), "NP_all"))

#DOC
DOC_TOTN_all <- bind_rows(mget(ls(pattern = "^DOC")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^DOC"), "DOC_TOTN_all"))

#Anions
Anions_all <- bind_rows(mget(ls(pattern = "^Anions")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Anions"), "Anions_all"))

#Cations
Cations_all <- bind_rows(mget(ls(pattern = "^Cations")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Cations"), "Cations_all"))

```

merge data - keep everything from TagTag, put NAs in the other locations
```{r}
TagTag <- TagTag %>%
  left_join(DOC_TOTN_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(NP_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Anions_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Cations_all, by = "SampleID")

#just for Asa that doesn't have the pH and EC in the same file as the TagTag
TagTag <- TagTag %>%
  left_join(EC_pH_Asa_2022_2025, by = "SampleID")

rm(DOC_TOTN_all, NP_all, Anions_all, Cations_all, EC_pH_Asa_2022_2025)
```

#Lake
##Data on Portal
Upload data - you need to download the most recent file on the portal for this
```{r}
skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHEM-L_ASA_FER_20170622-20251208_L2_irregular.csv"), "####"))[1]
OnPortal <- read_csv("OnPortal/SITES_CHEM-L_ASA_FER_20170622-20251208_L2_irregular.csv", skip = skip_lines)

#change LOD to error code 99999 and NaN to NA and convert everything to numeric 
OnPortal <- OnPortal %>%
  mutate(across(!any_of("TIMESTAMP"), as.character)) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == "LOD", "99999", .))) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ suppressWarnings(as.numeric(.)))) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(is.nan(.), NA_real_, .)))
```

##Data cleaning
**Just Asa - lake**
```{r}
#drop the cations 2017-08-19 outliers across the board 
keep_prefixes <- c("EC", "pH", "TW", "DOC", "TOT-N", "PO4-P", "NH4-N", "NO3-N", "F", "Cl", "Br", "SO4-S")

# Identify columns to nullify (excluding TIMESTAMP)
nullify_cols <- names(OnPortal)[!grepl(paste0("^(", paste(keep_prefixes, collapse = "|"), ")"), names(OnPortal))]
nullify_cols <- nullify_cols[nullify_cols != "TIMESTAMP"]

# Set values to NA for the target timestamp
OnPortal <- OnPortal %>%
  mutate(across(all_of(nullify_cols), ~ if_else(TIMESTAMP == as.POSIXct("2017-08-09 12:00:00", tz = "UTC"), NA_real_, .)))


#drop the SO4-S that is currently on the Portal between 2020-10-12 12:00:00 and 2022-09-05 12:00:00 -- this isn't needed anymore -- was for fixing bad data that was fixed in 2026
#so4_cols <- names(OnPortal)[grepl("^SO4-S", names(OnPortal))]
# OnPortal <- OnPortal %>%
#   mutate(across(all_of(so4_cols), ~ if_else(
#     TIMESTAMP >= as.POSIXct("2020-10-12 12:00:00", tz = "UTC") & 
#     TIMESTAMP <= as.POSIXct("2022-09-05 12:00:00", tz = "UTC"), 
#     NA_real_, .)))

#remove anions for 2018-04-23 12:00:00
anions <- names(OnPortal)[grepl("^(Cl|SO4-S)", names(OnPortal))]

OnPortal <- OnPortal %>%
  mutate(across(all_of(anions), ~ if_else(
    TIMESTAMP == as.POSIXct("2018-04-23 12:00:00", tz = "UTC"),
    NA_real_, .)))
```

##Reformat portal data
```{r}
#Pivot to long format
OnPortal_long <- OnPortal |> 
  pivot_longer(
    cols = -c(TIMESTAMP), 
    names_to = "Variable", 
    values_to = "Value")

#for locations with depth in the header
OnPortal_long <- OnPortal_long %>%
  mutate(D_WATER = as.numeric(str_extract(Variable, "-[0-9.]+")),
         Variable = str_remove(Variable, "_-[0-9.]+m?"))
OnPortal_long <- OnPortal_long %>%
  relocate(D_WATER, .after = Variable)

#Portal data to merge with the new data (need to keep the 99999 error code)
OnPortal_ToMerge <- OnPortal_long

#add a column saying that this is the data from the portal
OnPortal_long_ToPreview <-  OnPortal_long %>% 
  mutate(dataset = "OnPortal")

OnPortal_long_ToPreview <- OnPortal_long_ToPreview %>%
  mutate(Value = if_else(Value == 99999, NA_real_, Value))

#drop NAs
OnPortal_long_ToPreview <- OnPortal_long_ToPreview %>%
  drop_na(Value)
```

##Reformat new data for preview
filter out just the lake data based on the SAMPLE ID 
For Asa 31 is lake depth 0.5m 
        32 is lake depth 9m
```{r}
# add a depth column
NewData_long <- TagTag |> 
  filter(str_ends(SampleID, "31|32")) |> 
  mutate(D_WATER = case_when(
    str_ends(SampleID, "31") ~ -0.5,
    str_ends(SampleID, "32") ~ -9,
    TRUE ~ NA_real_))

NewData_long <- NewData_long |> 
  select(-SampleID)

NewData_long <- NewData_long |> 
  mutate(TIMESTAMP = as.POSIXct(paste(format(TIMESTAMP, "%Y-%m-%d"), "12:00:00"), format = "%Y-%m-%d %H:%M:%S"))

#Pivot to long format
NewData_long <- NewData_long |> 
  pivot_longer(
    cols = -c(TIMESTAMP, D_WATER), 
    names_to = "Variable", 
    values_to = "Value")

#Portal data to merge with the new data (need to keep the 99999 error code)
NewData_ToMerge <- NewData_long

#add a column saying that this is the new data
NewData_long_ToPreview <-  NewData_long %>% 
  mutate(dataset = "NewData")

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  relocate(D_WATER, .after = Variable)

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  mutate(Value = if_else(Value == 99999, NA_real_, Value))

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  drop_na(Value)
```

##Preview all data
```{r}
#combine the long formats of the old and new data here
chem_combined_long <- bind_rows(OnPortal_long_ToPreview, NewData_long_ToPreview)

library(ggforce) #for the qc plots

pdf(paste0("QC_plots/", StationName, "_", LakeSamplingLocation, "_QC_plots.pdf"), width = 11, height = 8.5)

n_pages <- nlevels(factor(chem_combined_long$Variable)) %/% 6 + 1 #if it divides evenly you don't need the plus one and it will throw and error but the error only means there is an extra page, which is fine. 

for (i in seq_len(n_pages)) {

p <- ggplot(chem_combined_long,
    aes(TIMESTAMP, Value, 
        color = factor(D_WATER), 
        shape = dataset)) +
    geom_line(aes(group = interaction(dataset, D_WATER)), linewidth = 0.3) +
    geom_point(size = 1) +
    facet_wrap_paginate(
      ~ Variable,
      scales = "free_y",
      ncol = 2,
      nrow = 3,
      page = i) +
    scale_color_manual(
      values = c(
        "-0.5" = "grey60",
        "-9" = "black"),
      name = "Depth (m)") +
    scale_shape_manual(
      values = c(
        "OnPortal" = 4,   # x
        "NewData" = 21),  # closed dot=16
      name = "Dataset") +
    labs(
      title = paste("QC plots — page", i),
      x = "Date",
      y = "Value") +
    theme_light()
  print(p)   # one page per iteration
}

dev.off()

```

##Format Lake Data for upload
```{r}
OnPortal_ToMerge <- OnPortal_ToMerge |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC"))

NewData_ToMerge <- NewData_ToMerge |> 
  mutate(TIMESTAMP = as.POSIXct(paste(format(TIMESTAMP, "%Y-%m-%d"), "12:00:00"), format = "%Y-%m-%d %H:%M:%S"))

NewData_ToMerge <- NewData_ToMerge |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC"))

#distinct() keeps the first occurrence, so values from OnPortal_ToMerge are prioritized since it's listed first in bind_rows().
#If you want NewData_ToMerge to take priority, reverse the order:
LakeChemistry <- bind_rows(OnPortal_ToMerge, NewData_ToMerge) %>%
  distinct(TIMESTAMP, Variable, D_WATER, .keep_all = TRUE)

LakeChemistry <- LakeChemistry |> 
  pivot_wider(
    names_from = c(Variable, D_WATER),
    values_from = Value,
    names_glue = "{Variable}_{D_WATER}m"
  )

#this is what is in the file that is currently on the portal 
variable_order <- c("TIMESTAMP","EC_-0.5m","EC_-9m","pH_-0.5m","pH_-9m","TW_pH_-0.5m","TW_pH_-9m","DOC_-0.5m","DOC_-9m","TOT-N_-0.5m","TOT-N_-9m","PO4-P_-0.5m","PO4-P_-9m","NH4-N_-0.5m","NH4-N_-9m","NO3-N_-0.5m","NO3-N_-9m","F_-0.5m","F_-9m","Cl_-0.5m","Cl_-9m","Br_-0.5m","Br_-9m","SO4-S_-0.5m","SO4-S_-9m","Li_-0.5m","Li_-9m","Be_-0.5m","Be_-9m","B_-0.5m","B_-9m","Na_-0.5m","Na_-9m","Mg_-0.5m","Mg_-9m","Al_-0.5m","Al_-9m","Si_-0.5m","Si_-9m","P_-0.5m","P_-9m","S_-0.5m","S_-9m","K_-0.5m","K_-9m","Ca_-0.5m","Ca_-9m","V_-0.5m","V_-9m","Cr_-0.5m","Cr_-9m","Mn_-0.5m","Mn_-9m","Fe_-0.5m","Fe_-9m","Co_-0.5m","Co_-9m","Ni_-0.5m","Ni_-9m","Cu_-0.5m","Cu_-9m","Zn_-0.5m","Zn_-9m","Ga_-0.5m","Ga_-9m","As_-0.5m","As_-9m","Se_-0.5m","Se_-9m","Rb_-0.5m","Rb_-9m","Sr_-0.5m","Sr_-9m","Ag_-0.5m","Ag_-9m","Cd_-0.5m","Cd_-9m","Te_-0.5m","Te_-9m","Cs_-0.5m","Cs_-9m","Ba_-0.5m","Ba_-9m","Tl_-0.5m","Tl_-9m","Pb_-0.5m","Pb_-9m","Bi_-0.5m","Bi_-9m")

LakeChemistry <- LakeChemistry |>  
  select(all_of(variable_order)) |>
  mutate(across(-TIMESTAMP, ~ suppressWarnings(round(as.numeric(.), 4))),
         across(-TIMESTAMP, ~ ifelse(. == 99999, "LOD", .)),
         across(-TIMESTAMP, ~ ifelse(is.na(.), NaN, .)))

LakeChemistry <- LakeChemistry %>% 
mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))
```

##Export Lake data for upload
```{r}
# Construct the filename
filename <- paste0("FinalChemFiles/CHEM-L_", StationName, "_", LakeSamplingLocation, "_","until" ,EndYear, ".csv")

#Convert all to character so that you can maintain the NaN
LakeChemistry_export <- LakeChemistry %>%
  mutate(across(-TIMESTAMP, ~ ifelse(is.nan(.), "NaN", as.character(.))))

# Export to CSV
write.csv(LakeChemistry_export, filename, row.names = FALSE, quote = FALSE)
```

##Clear lake files
```{r}
rm(chem_combined_long, LakeChemistry, LakeChemistry_export,
   OnPortal, OnPortal_long, OnPortal_long_ToPreview, OnPortal_ToMerge,
   NewData_long, NewData_long_ToPreview, NewData_ToMerge)

rm(LakeSamplingLocation, filename, df_names, df_name)
```


#Stream
If you add a lot of additional stream sampling points you might want to split this out and do one preview by location -- right now we are going to keep them together
```{r}
StreamSamplingLocation <- "STREAMS"
```

##Data on Portal
Upload data - you need to download the most recent file on the portal for this
```{r}
#Outlet
skip_lines_OUT <- which(str_detect(readLines("OnPortal/SITES_CHEM-S_ASA_OUT_20170622-20251208_L2_irregular.csv"), "####"))[1]
OnPortal_OUT <- read_csv("OnPortal/SITES_CHEM-S_ASA_OUT_20170622-20251208_L2_irregular.csv", skip = skip_lines_OUT)

#change LOD to error code 99999 and NaN to NA and convert everything to numeric 
OnPortal_OUT <- OnPortal_OUT |> 
  mutate(across(!any_of("TIMESTAMP"), as.character)) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == "LOD", "99999", .))) |> 
  mutate(across(!any_of("TIMESTAMP"), ~ suppressWarnings(as.numeric(.)))) |> 
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(is.nan(.), NA_real_, .)))

#Main Inlet
skip_lines_MIN <- which(str_detect(readLines("OnPortal/SITES_CHEM-S_ASA_MIN_20170622-20251208_L2_irregular.csv"), "####"))[1]
OnPortal_MIN <- read_csv("OnPortal/SITES_CHEM-S_ASA_MIN_20170622-20251208_L2_irregular.csv", skip = skip_lines_MIN)

#change LOD to error code 99999 and NaN to NA and convert everything to numeric 
OnPortal_MIN <- OnPortal_MIN |> 
  mutate(across(!any_of("TIMESTAMP"), as.character)) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == "LOD", "99999", .))) |> 
  mutate(across(!any_of("TIMESTAMP"), ~ suppressWarnings(as.numeric(.)))) |> 
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(is.nan(.), NA_real_, .)))

```

##Data cleaning - old data
*Specific cleaning for Asa streams*
```{r}
#This is no longer needed -- was done during the 2026 update to deal with old incorrect data on the portal
#drop the SO4-S that is currently on the Portal between 2020-10-12 12:00:00 and 2022-09-05 12:00:00 - this is the old stat Christoffer didn't convert
# OnPortal_OUT <- OnPortal_OUT |> 
#   mutate(`SO4-S` = if_else(
#     TIMESTAMP >= as.POSIXct("2020-10-12 12:00:00", tz = "UTC") & 
#     TIMESTAMP <= as.POSIXct("2022-09-05 12:00:00", tz = "UTC"), 
#     NA_real_, `SO4-S`))
# OnPortal_MIN <- OnPortal_MIN |> 
#   mutate(`SO4-S` = if_else(
#     TIMESTAMP >= as.POSIXct("2020-10-12 12:00:00", tz = "UTC") & 
#     TIMESTAMP <= as.POSIXct("2022-09-05 12:00:00", tz = "UTC"), 
#     NA_real_, `SO4-S`))


#Should this outlier be removed? Other points with the same timestamp aren't weird
#remove Ni outlier - OUT ? 2021-08-09 12:00:00
# OnPortal_OUT <- OnPortal_OUT |> 
#   mutate(`Ni` = if_else(
#     TIMESTAMP == as.POSIXct("2021-08-09 12:00:00", tz = "UTC"),
#     NA_real_, `Ni`))
```

##Reformat portal data
```{r}
#Pivot to long format
##Outlet data
OnPortal_OUT_long <- OnPortal_OUT |> 
  pivot_longer(
    cols = -c(TIMESTAMP), 
    names_to = "Variable", 
    values_to = "Value")

#add the sampling point
OnPortal_OUT_long <- OnPortal_OUT_long %>%
  mutate(SamplingPoint = "OUT")

#Portal data to merge with the new data (need to keep the 99999 error code)
OnPortal_OUT_ToMerge <- OnPortal_OUT_long

#On Portal data to preview - add a column saying that this is the data from the portal
OnPortal_OUT_long_ToPreview <-  OnPortal_OUT_long %>% 
  mutate(dataset = "OnPortal")

OnPortal_OUT_long_ToPreview <- OnPortal_OUT_long_ToPreview %>%
  mutate(Value = if_else(Value == 99999, NA_real_, Value))

#drop NAs
OnPortal_OUT_long_ToPreview <- OnPortal_OUT_long_ToPreview %>%
  drop_na(Value)

#Inlet
OnPortal_MIN_long <- OnPortal_MIN |> 
  pivot_longer(
    cols = -c(TIMESTAMP), 
    names_to = "Variable", 
    values_to = "Value")

#add the sampling point
OnPortal_MIN_long <- OnPortal_MIN_long %>%
  mutate(SamplingPoint = "MIN")

#Portal data to merge with the new data (need to keep the 99999 error code)
OnPortal_MIN_ToMerge <- OnPortal_MIN_long

#add a column saying that this is the data from the portal
OnPortal_MIN_long_ToPreview <-  OnPortal_MIN_long %>% 
  mutate(dataset = "OnPortal")

OnPortal_MIN_long_ToPreview <- OnPortal_MIN_long_ToPreview %>%
  mutate(Value = if_else(Value == 99999, NA_real_, Value))

#drop NAs
OnPortal_MIN_long_ToPreview <- OnPortal_MIN_long_ToPreview %>%
  drop_na(Value)

```

##Select stream data
filter out just the strea data based on the SAMPLE ID 
For Asa 11 is stream inlet
        21 is stream outlet
```{r}
# add a depth column
NewData_long <- TagTag |> 
  filter(str_ends(SampleID, "11|21")) |> 
  mutate(SamplingPoint = case_when(
    str_ends(SampleID, "11") ~ "MIN",
    str_ends(SampleID, "21") ~ "OUT",
    TRUE ~ NA_character_))

NewData_long <- NewData_long |> 
  select(-SampleID)

NewData_long <- NewData_long |> 
  mutate(TIMESTAMP = as.POSIXct(paste(format(TIMESTAMP, "%Y-%m-%d"), "12:00:00"), format = "%Y-%m-%d %H:%M:%S"))
```

##Data cleaning - new data
*Just Asa Streams* 
Should this outlier be removed? Other points with the same timestamp aren't weird
```{r}
# NewData_long <- NewData_long |>
#   mutate(`Ni` = if_else(
#     TIMESTAMP == as.POSIXct("2021-08-09 12:00:00") &
#     SamplingPoint == "OUT", 
#     NA_real_, `Ni`))
```

##Reformat new data for preview
```{r}
#Pivot to long format
NewData_long <- NewData_long |> 
  pivot_longer(
    cols = -c(TIMESTAMP, SamplingPoint), 
    names_to = "Variable", 
    values_to = "Value")

#Portal data to merge with the new data (need to keep the 99999 error code)
NewData_ToMerge <- NewData_long

#add a column saying that this is the new data
NewData_long_ToPreview <-  NewData_long %>% 
  mutate(dataset = "NewData")

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  relocate(SamplingPoint, .after = Value)

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  mutate(Value = if_else(Value == 99999, NA_real_, Value))

NewData_long_ToPreview <- NewData_long_ToPreview %>%
  drop_na(Value)
```

##Preview all data
```{r}
#combine the long formats of the old and new data here
stream_chem_combined_long <- bind_rows(OnPortal_MIN_long_ToPreview, OnPortal_OUT_long_ToPreview, NewData_long_ToPreview)

library(ggforce) #for the qc plots

pdf(paste0("QC_plots/", StationName, "_", StreamSamplingLocation, "_QC_plots.pdf"), width = 11, height = 8.5)

n_pages <- nlevels(factor(stream_chem_combined_long$Variable)) %/% 6 + 1 #if it divides evenly you don't need the plus one and it will throw and error but the error only means there is an extra page, which is fine. 

for (i in seq_len(n_pages)) {

p <- ggplot(stream_chem_combined_long,
    aes(TIMESTAMP, Value, 
        color = factor(SamplingPoint), 
        shape = dataset)) +
    geom_line(aes(group = interaction(dataset, SamplingPoint)), linewidth = 0.3) +
    geom_point(size = 1) +
    facet_wrap_paginate(
      ~ Variable,
      scales = "free_y",
      ncol = 2,
      nrow = 3,
      page = i) +
    scale_color_manual(
      values = c(
        "MIN" = "grey60",
        "OUT" = "black"),
      name = "Sampling Point") +
    scale_shape_manual(
      values = c(
        "OnPortal" = 4,   # x
        "NewData" = 21),  # closed dot
      name = "Dataset") +
    labs(
      title = paste("QC plots — page", i),
      x = "Date",
      y = "Value") +
    theme_light()
  print(p)   # one page per iteration
}

dev.off()

```

Drop the dfs used to preview the data
```{r}
rm(OnPortal_MIN, OnPortal_MIN_long_ToPreview, OnPortal_MIN_long, 
   OnPortal_OUT, OnPortal_OUT_long_ToPreview, OnPortal_OUT_long, 
   NewData_long_ToPreview, NewData_long, 
   stream_chem_combined_long)
```


##Format and print Stream Data for upload
###OUT
```{r}
# Specify sampling point here
sampling_point <- "OUT"

#You are dropping the NAs here because when you reformat it to wide you will replace the NAs with NaN
OnPortal_SP_ToMerge <- get(paste0("OnPortal_", sampling_point, "_ToMerge")) |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC")) |> 
  select(-SamplingPoint)

#specify which new data
NewData_SP_ToMerge <- NewData_ToMerge |> 
  filter(SamplingPoint == sampling_point)|> 
  select(-SamplingPoint)
NewData_SP_ToMerge <- NewData_SP_ToMerge |> 
  mutate(TIMESTAMP = as.POSIXct(paste(format(TIMESTAMP, "%Y-%m-%d"), "12:00:00"), format = "%Y-%m-%d %H:%M:%S"))
NewData_SP_ToMerge <- NewData_SP_ToMerge |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC"))

#distinct() keeps the first occurrence, so values from OnPortal_SP_ToMerge are prioritized since it's listed first in bind_rows().
#If you want NewData_SP_ToMerge to take priority, reverse the order:
StreamChemistry_SP <- bind_rows(OnPortal_SP_ToMerge, NewData_SP_ToMerge) %>%
  distinct(TIMESTAMP, Variable, .keep_all = TRUE)
StreamChemistry_SP <- StreamChemistry_SP |> 
  pivot_wider(
    names_from = c(Variable),
    values_from = Value)

#this is what is in the file that is currently on the portal 
variable_order <- c("TIMESTAMP","EC","pH","TW_pH","DOC","TOT-N","PO4-P","NH4-N","NO3-N","F","Cl","Br","SO4-S","Li","Be","B","Na","Mg",
                    "Al","Si","P","S","K","Ca","V","Cr","Mn","Fe","Co","Ni","Cu","Zn","Ga","As","Se","Rb","Sr","Ag","Cd","Te","Cs","Ba",
                    "Tl","Pb","Bi")
StreamChemistry_SP <- StreamChemistry_SP |>  
  select(all_of(variable_order)) |>
  mutate(across(-TIMESTAMP, ~ suppressWarnings(round(as.numeric(.), 4))),
         across(-TIMESTAMP, ~ ifelse(. == 99999, "LOD", .)),
         across(-TIMESTAMP, ~ ifelse(is.na(.), NaN, .)))
StreamChemistry_SP <- StreamChemistry_SP %>% 
  mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))

##Export##
# Construct the filename
filename <- paste0("FinalChemFiles/CHEM-S_", StationName, "_", sampling_point, "_","until" ,EndYear, ".csv")

#Convert all to character so that you can maintain the NaN
StreamChemistry_export <- StreamChemistry_SP %>%
  mutate(across(-TIMESTAMP, ~ ifelse(is.nan(.), "NaN", as.character(.))))

# Export to CSV
write.csv(StreamChemistry_export, filename, row.names = FALSE, quote = FALSE)
```

drop dfs
```{r}
rm(OnPortal_OUT_ToMerge, OnPortal_SP_ToMerge, 
   NewData_SP_ToMerge,
   StreamChemistry_SP, filename, StreamChemistry_export)
```


###MIN
```{r}
# Specify sampling point here
sampling_point <- "MIN"

#You are dropping the NAs here because when you reformat it to wide you will replace the NAs with NaN
OnPortal_SP_ToMerge <- get(paste0("OnPortal_", sampling_point, "_ToMerge")) |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC")) |> 
  select(-SamplingPoint)

#specify which new data
NewData_SP_ToMerge <- NewData_ToMerge |> 
  filter(SamplingPoint == sampling_point)|> 
  select(-SamplingPoint)
NewData_SP_ToMerge <- NewData_SP_ToMerge |> 
  mutate(TIMESTAMP = as.POSIXct(paste(format(TIMESTAMP, "%Y-%m-%d"), "12:00:00"), format = "%Y-%m-%d %H:%M:%S"))
NewData_SP_ToMerge <- NewData_SP_ToMerge |> 
  drop_na(Value) |> 
  mutate(TIMESTAMP = force_tz(TIMESTAMP, tzone = "UTC"))

#distinct() keeps the first occurrence, so values from OnPortal_SP_ToMerge are prioritized since it's listed first in bind_rows().
#If you want NewData_SP_ToMerge to take priority, reverse the order:
StreamChemistry_SP <- bind_rows(OnPortal_SP_ToMerge, NewData_SP_ToMerge) %>%
  distinct(TIMESTAMP, Variable, .keep_all = TRUE)
StreamChemistry_SP <- StreamChemistry_SP |> 
  pivot_wider(
    names_from = c(Variable),
    values_from = Value)

#this is what is in the file that is currently on the portal 
variable_order <- c("TIMESTAMP","EC","pH","TW_pH","DOC","TOT-N","PO4-P","NH4-N","NO3-N","F","Cl","Br","SO4-S","Li","Be","B","Na","Mg",
                    "Al","Si","P","S","K","Ca","V","Cr","Mn","Fe","Co","Ni","Cu","Zn","Ga","As","Se","Rb","Sr","Ag","Cd","Te","Cs","Ba",
                    "Tl","Pb","Bi")
StreamChemistry_SP <- StreamChemistry_SP |>  
  select(all_of(variable_order)) |>
  mutate(across(-TIMESTAMP, ~ suppressWarnings(round(as.numeric(.), 4))),
         across(-TIMESTAMP, ~ ifelse(. == 99999, "LOD", .)),
         across(-TIMESTAMP, ~ ifelse(is.na(.), NaN, .)))
StreamChemistry_SP <- StreamChemistry_SP %>% 
  mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))

##Export##
# Construct the filename
filename <- paste0("FinalChemFiles/CHEM-S_", StationName, "_", sampling_point, "_","until" ,EndYear, ".csv")

#Convert all to character so that you can maintain the NaN
StreamChemistry_export <- StreamChemistry_SP %>%
  mutate(across(-TIMESTAMP, ~ ifelse(is.nan(.), "NaN", as.character(.))))

# Export to CSV
write.csv(StreamChemistry_export, filename, row.names = FALSE, quote = FALSE)
```

drop dfs
```{r}
rm(OnPortal_MIN_ToMerge, OnPortal_SP_ToMerge, 
   NewData_SP_ToMerge,
   StreamChemistry_SP, filename, StreamChemistry_export)
```


