---
title: "Cations"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

load packages
```{r}
library(tidyverse)
library(readxl)
```


#GU - lab

load data - sometimes the filenames are different so you cant fully automate
*CHANGE SOMETHING MANUALLY HERE*
```{r}
rawfile <- "cations_GU_Skogaryd_23" #ENTER YOUR FILENAME HERE

#use for .csv
filepath <- paste0("LabResults/Cations/", rawfile, ".csv")
df <- read_csv(filepath)

#use for .xlsx
filepath <- paste0("LabResults/Cations/", rawfile, ".xlsx")
df <- read_excel(filepath)

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "Skogaryd" #STATION NAME HERE
StartYear <- 2023 #START YEAR HERE
EndYear <- 2023 #END YEAR HERE
```

clean up
```{r}
#drop first column
df <- df |> 
  select(-`Measurement session`, -any_of("element")) #this only removes the column element if it exists (inconsistent naming from lab)

df <- df %>%
  drop_na(`Sample Name`)

#only keep the columns that don't start with ... (those had RSD in them)
df <- df %>%
  select(!starts_with("..."))
```

Double check that you don't have any 9999s
This returns TRUE if any cell equals 9999, and FALSE if no cells equal 9999.
```{r}
any(df == 9999, na.rm = TRUE)
```

*as long as the answer above is FALSE continue - if not you need more code*
deal with LOD
We are going to change anything with a < to to 9999 and then change that to LOD after doing the unit conversion
```{r}
df <- df %>%
  mutate(across(everything(), ~ ifelse(grepl("^<", .), 9999, .)))

df <- df %>%
  mutate(across(-1, as.numeric))

df <- df %>%
  mutate(across(-`Sample Name`, ~ ifelse(. == 9999, "LOD", .)))
```

Rename
```{r}
df <- df %>%
  rename(
    SampleID = `Sample Name`)
```

Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/", StationName, "/Cations_", StationName, "_", StartYear,"_",EndYear, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE, quote = FALSE)
```




