---
title: "DOC and N-TOT"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

NOTE - there is different code for the results that come from Skogaryd and the results from SLU Umeå lab

load packages
```{r}
library(tidyverse)
library(readxl)
```

#SLU Umeå lab
load data - sometimes the filenames are different so you cant fully automate
Make sure that in the raw data in the RESULTS tab you drop the columns with RSD

**CHANGE SOMETHING MANUALLY HERE**
```{r}
rawfile <- "TOC_TN_Umeå_Asa_2025" #ENTER YOUR FILENAME HERE

filepath <- paste0("LabResults/TOC/", rawfile, ".xlsx")
df <- read_excel(filepath, sheet = "RESULTS")

#WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA 
StationName <- "Asa" #STATION NAME HERE
StartYear <- 2025 #START YEAR HERE
EndYear <- 2026 #END YEAR HERE
```

**check results** - if something other than mg/L shows up you will need to do a conversion
```{r}
unique(as.character(unlist(df[1, ])))
```

as long as the above is mg/L - continue
```{r}
#drop the first row
df <- df %>%
  slice(-1)

#drop unnecessary columns
df <- df %>%
  select(-any_of(c("CUSTID 1", "CUSTID 2", "KUNDID 1", "KUNDID", "Cust ID 1", "Cust ID 2")))
```

**check data**
Check whether there are any that = 9999 (using this as the error code)
```{r}
df %>%
  summarise(across(everything(), ~ any(. == 9999, na.rm = TRUE)))
```

**check results**
Check whether there are any LODs in the results - if there are you need to convert them to 9999 and then convert them back later 
```{r}
df %>%
  summarise(across(everything(), ~ any(grepl("^<$|^LOD$|^<LOD$", ., ignore.case = TRUE), na.rm = TRUE))) #this shows anything that has < or LOD or <LOD, this should also catch any weird spaces in the LOD cell

#only need to run this if there are LOD above
# df <- df %>%
#   mutate(across(everything(), ~ ifelse(grepl("^<|^LOD$", ., ignore.case = TRUE), 9999, .)))
```

convert to numeric
```{r}
df <- df %>%  # convert all columns except the first to numeric
 mutate(across(-1, as.numeric))
```

Rename
Note - there is no LOD for this data
```{r}
df <- df %>%
  rename(
    SampleID = `Lab ID`,
    DOC = `NPOC`,
    `TOT-N` = `TN`)
```

Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/", StationName, "/DOC_TOTN_", StationName, "_", StartYear,"_",EndYear, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE, quote = FALSE)
```


#Skogaryd lab
They run this themselves and the format is specific to them
load data

NOTE - 2026-06-25 -- there is one date in 2025 where it seems like data should be re-calcualted but it wasn't, that needs to be added to the raw file and then things can be run through here. 
2024 - is missing all but 1 file - e-mailed Aila to get access to the rest

*CHANGE SOMETHING MANUALLY HERE*
```{r}
#formatted for excel files
rawfile <- "TOC_GU_Skogaryd_2025" #ENTER YOUR FILENAME HERE

filepath <- paste0("LabResults/TOC/", rawfile, ".xlsx")
df <- read_excel(filepath, sheet = "RESULTS")

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "Skogaryd" #STATION NAME HERE
StartYear <- 2025 #START YEAR HERE
EndYear <- 2025 #END YEAR HERE
```

drop unnecessary columns and rows
```{r}
#drop anything above the data
df <- df %>%
  slice(-(1:which(.[[1]] == "[Data]")[1]))

df <- df %>%
  slice(-1) %>%              # Remove first row after storing names
  setNames(as.character(df[1, ]))  # Set first row as names

df <- df %>%
  select(-any_of(c("Type", "Sample ID", "Vial", "Date / Time")))

```

*check results*
```{r}
#units - we use mg/L - double check that it's only mg/L displayed when you run this
df %>%
  distinct(Unit) %>%
  pull(Unit)
```

more cleaning
```{r}
#as long as it's only mg/L above you can drop this column
df <- df %>%
  select(-("Unit"))

#remove the things that are blanks etc. just keep Skogaryd data (5-)
df <- df %>%
  filter(grepl("^5-", df[[1]]))

df <- df %>%  # convert all columns except the first to numeric
 mutate(across(-1, as.numeric))

#change the number of decimals - different for the two columns
df <- df %>%
  mutate(`Result(NPOC)` = round(as.numeric(`Result(NPOC)`), 2)) %>%
  mutate(`Result(TN)` = round(as.numeric(`Result(TN)`), 4))
```

Rename
Note - there is no LOD for this data
```{r}
df <- df %>%
  rename(
    SampleID = `Sample Name`,
    DOC = `Result(NPOC)`,
    `TOT-N` = `Result(TN)`)
```


Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/", StationName, "/DOC_TOTN_", StationName, "_", StartYear,"_",EndYear, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE, quote = FALSE)
```







