---
title: "Merge Stream Chemistry Datasets"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

load packages
```{r}
library(tidyverse)
```

load data
```{r}
#Example
# TagTag <- read_csv("ProcessedData/TagTag_Skogaryd_2024.csv")
# Anions_Skogaryd_2023_2023 <- read_csv("ProcessedData/Anions_Skogaryd_2023-2023.csv")

#Add you data here - you could maybe change this so that it uploads an entire folder
TagTag <- read_csv("ProcessedData/TagTag_Abisko_Miellejohka_2022.csv")

NP_Abisko_2021_2022 <- read_csv("ProcessedData/NP_Abisko_2021_2022.csv")
NP_Abisko_2022_2023 <- read_csv("ProcessedData/NP_Abisko_2022_2023.csv")
NP_Abisko_2023_2023 <- read_csv("ProcessedData/NP_Abisko_2023_2023.csv")
NP_Abisko_2023_2024 <- read_csv("ProcessedData/NP_Abisko_2023_2024.csv")

DOC_TOTN_Abisko_2022_2022 <- read_csv("ProcessedData/DOC_NTOT_Abisko_2022_2022.csv")
DOC_TOTN_Abisko_2022_2023 <- read_csv("ProcessedData/DOC_NTOT_Abisko_2022_2023.csv")
DOC_TOTN_Abisko_2023_2023 <- read_csv("ProcessedData/DOC_NTOT_Abisko_2023_2023.csv")
DOC_TOTN_Abisko_2023_2024 <- read_csv("ProcessedData/DOC_NTOT_Abisko_2023_2024.csv")

Anions_Abisko_2021_2022 <- read_csv("ProcessedData/Anions_Abisko_2021_2022.csv")
Anions_Abisko_2022_2023 <- read_csv("ProcessedData/Anions_Abisko_2022_2023.csv")
Anions_Abisko_2023_2024 <- read_csv("ProcessedData/Anions_Abisko_2023_2024.csv")

Cations_Abisko_2021_2022 <- read_csv("ProcessedData/Cations_Abisko_2021_2022.csv")
Cations_Abisko_2022_2024 <- read_csv("ProcessedData/Cations_Abisko_2022_2024.csv")
Cations_Abisko_2024_2024 <- read_csv("ProcessedData/Cations_Abisko_2024_2024.csv")
```

```{r}
# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "ANS" #STATION Abrevitation here
SamplingLocation <- "MJH" #sampling code here
EndYear <- 2024 #END YEAR HERE
```


change everything to character since you are merging some that have LOD
```{r}
TagTag <- TagTag %>%
  mutate(across(-TIMESTAMP, as.character))

# Get all dataframe names in the environment except TagTag
df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
df_names <- df_names[df_names != "TagTag"]

# Convert all columns to character for each dataframe
for(df_name in df_names) {
  assign(df_name, get(df_name) %>% mutate(across(everything(), as.character)))
}
```

merge any overlapping datasets
```{r}
#NP
NP <- bind_rows(NP_Abisko_2021_2022, NP_Abisko_2022_2023, NP_Abisko_2023_2023, NP_Abisko_2023_2024)

rm(NP_Abisko_2021_2022, NP_Abisko_2022_2023, NP_Abisko_2023_2023, NP_Abisko_2023_2024)

#DOC
DOC_TOTN <- bind_rows(DOC_TOTN_Abisko_2022_2022,DOC_TOTN_Abisko_2022_2023, DOC_TOTN_Abisko_2023_2023, DOC_TOTN_Abisko_2023_2024)

rm(DOC_TOTN_Abisko_2022_2022, DOC_TOTN_Abisko_2022_2023, DOC_TOTN_Abisko_2023_2023, DOC_TOTN_Abisko_2023_2024)

#Anions
Anions <- bind_rows(Anions_Abisko_2021_2022, Anions_Abisko_2022_2023, Anions_Abisko_2023_2024)

rm(Anions_Abisko_2021_2022, Anions_Abisko_2022_2023, Anions_Abisko_2023_2024)


#Cations
Cations <- bind_rows(Cations_Abisko_2021_2022 , Cations_Abisko_2022_2024, Cations_Abisko_2024_2024)

rm(Cations_Abisko_2021_2022, Cations_Abisko_2022_2024, Cations_Abisko_2024_2024)

```


merge data - keep everything from TagTag leave NAs in the other locations
```{r}
TagTag <- TagTag %>%
  left_join(DOC_TOTN, by = "SampleID")

TagTag <- TagTag %>%
  left_join(NP, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Anions, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Cations, by = "SampleID")

rm(DOC_TOTN, NP, Anions, Cations)

#just for abisko - drop F until things are re-run - there was a problem with a difference in magnitude between the two labs
TagTag <- TagTag %>% 
  select(-"F")
```


#Data on Portal
##Upload data
Almbergasjön: SITES_CHEM-L_ANS_ALP_20170626-20220924_L2_irregular
Miellejohka: SITES_CHEM-S_ANS_MJH_20170407-20220920_L2_irregular
```{r}
lines <- readLines("OnPortal/SITES_CHEM-S_ANS_MJH_20170407-20220920_L2_irregular.csv")
skip_lines <- which(str_detect(lines, "####"))[1]
OnPortal <- read_csv("OnPortal/SITES_CHEM-S_ANS_MJH_20170407-20220920_L2_irregular.csv", skip = skip_lines)

#just for abisko 
OnPortal <- OnPortal %>%
  filter(TIMESTAMP <= as.Date("2022-04-05"))
```

##Reformat for preview
```{r}
#change LOD to NA for the preview
OnPortal_long <- OnPortal %>%
  mutate(across(where(is.character) & -TIMESTAMP, ~na_if(., "LOD")))


#change NaN to NA and convert to numeric for the preview
OnPortal_long <- OnPortal_long %>%
  mutate(across(where(~is.numeric(.) | is.character(.)) & -TIMESTAMP, 
                ~suppressWarnings(as.numeric(ifelse(. == "NaN", NA, .)))))

#Pivot to long format
OnPortal_long <- OnPortal_long |> 
  pivot_longer(cols = -c(TIMESTAMP), names_to = "Variable", values_to = "Value")

OnPortal_long <- OnPortal_long %>%
  drop_na(Value)

#add a column saying that this is the data from the portal
OnPortal_long <-  OnPortal_long %>% 
  mutate(dataset = "OnPortal")
```

reformat new data for preview
```{r}
TagTag_long <- TagTag %>% 
  select(-SampleID)

TagTag_long <- TagTag_long %>%
  mutate(TIMESTAMP = as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"))

#change LOD to NA for the preview
TagTag_long <- TagTag_long %>%
  mutate(across(where(is.character) & -TIMESTAMP, ~na_if(., "LOD")))

#change NaN to NA and convert to numeric for the preview
TagTag_long <- TagTag_long %>%
  mutate(across(where(~is.numeric(.) | is.character(.)) & -TIMESTAMP, 
                ~suppressWarnings(as.numeric(ifelse(. == "NaN", NA, .)))))

#Pivot to long format
TagTag_long <- TagTag_long |> 
  pivot_longer(cols = -c(TIMESTAMP), names_to = "Variable", values_to = "Value")

TagTag_long <- TagTag_long %>%
  drop_na(Value)

#add a column saying that this is the new data
TagTag_long <-  TagTag_long %>% 
  mutate(dataset = "NewData")
```


#Preview all data
NOTE - change Date to TIMESTAMP in the ggplot code when Robyn sends you the sampling protocols from 2023
```{r}
#combine the long formats of the old and new data here
chem_combined_long <- bind_rows(OnPortal_long, TagTag_long)

library(ggforce) #for the qc plots

pdf(paste0(StationName, "_", SamplingLocation, "_QC_plots.pdf"), width = 11, height = 8.5)

n_pages <- nlevels(factor(chem_combined_long$Variable)) %/% 6 + 1 #if it divides evenly you don't need the plus one and it will throw and error but the error only means there is an extra page, which is fine. 

for (i in seq_len(n_pages)) {

  p <- ggplot(chem_combined_long,
    aes(TIMESTAMP, Value, color = dataset)) +
    geom_line(linewidth = 0.3) +
    geom_point(size = 0.6) +
    facet_wrap_paginate(
      ~ Variable,
      scales = "free_y",
      ncol = 2,
      nrow = 3,
      page = i) +
    scale_color_manual(values = c(
      "OnPortal" = "grey60",
      "NewData"  = "black")) +
    labs(
      title = paste("QC plots — page", i),
      x = "Date",
      y = "Value",
      color = "Dataset") +
    theme_light()

  print(p)   # one page per iteration
}

dev.off()

```

remove the long format data
```{r}
rm(OnPortal_long, TagTag_long, chem_combined_long)
```


#Merge the old and new data for export
```{r}
variable_order <- c("TIMESTAMP","pH","DOC","TOT-N","PO4-P","NH4-N","NO3-N","Cl","SO4-S","Br","Li","Be","B","Na","Mg","Al","Si","P","S","K","Ca","V","Cr","Mn","Fe","Co","Ni","Cu","Zn","Ga","As","Se","Rb","Sr","Ag","Cd","In","Te","Cs","Ba","Tl","Pb","Bi")

TagTag <- TagTag %>% 
  select(-SampleID) %>%
  mutate(`SO4-S` = as.character(round(as.numeric(`SO4-S`), 4))) %>%
  select(all_of(variable_order)) %>%
  mutate(across(-TIMESTAMP, ~replace(., is.na(.), NaN)))

OnPortal <- OnPortal %>%
  mutate(across(-1, as.character))%>%
  select(all_of(variable_order))

Stream_chemistry <- bind_rows(OnPortal, TagTag)

Stream_chemistry <- Stream_chemistry %>% 
mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))
```


```{r}
# Construct the filename
filename <- paste0("FinalChemFiles/CHEM-S", StationName, "_", SamplingLocation, "_","until" ,EndYear, ".csv")

# Export to CSV
write.csv(Stream_chemistry, filename, row.names = FALSE, quote = FALSE)
```

