---
title: "Merge Lake Chemistry Datasets"
author: "Matilda"
date: "2026-01-30"
output: html_document
---


load packages
```{r}
library(tidyverse)
library(readxl)
```

load data
```{r}
#Example
# TagTag <- read_csv("ProcessedData/TagTag_Skogaryd_2024.csv")
# Anions_Skogaryd_2023_2023 <- read_csv("ProcessedData/Anions_Skogaryd_2023-2023.csv")

#Add you data here - you could maybe change this so that it uploads an entire folder
#TagTag 
#TagTag <- read_csv("ProcessedData/FILENAME.csv")
#or SampleID file
TagTag <- read_excel("TagTag/ASA/TagTag_Asa.xlsx")

#Anions
Anions_Asa_2020_2021 <- read_csv("ProcessedData/Anions_Asa_2020_2021.csv")
Anions_Asa_2021_2022 <- read_csv("ProcessedData/Anions_Asa_2021_2022.csv")
Anions_Asa_2022_2023 <- read_csv("ProcessedData/Anions_Asa_2022_2023.csv")
Anions_Asa_2024_2024 <- read_csv("ProcessedData/Anions_Asa_2024_2024.csv")

#Cations
Cations_Asa_2017_2022 <- read_csv("ProcessedData/Cations_Asa_2017_2022.csv")
Cations_Asa_2022_2023 <- read_csv("ProcessedData/Cations_Asa_2022_2023.csv")
Cations_Asa_2023_2024 <- read_csv("ProcessedData/Cations_Asa_2023_2024.csv")

#DOC and TOTN
DOC_TOTN_Asa_2022_2022 <- read_csv("ProcessedData/DOC_TOTN_Asa_2022_2022.csv")
DOC_TOTN_Asa_2023_2023 <- read_csv("ProcessedData/DOC_TOTN_Asa_2023_2023.csv")
DOC_TOTN_Asa_2024_2024 <- read_csv("ProcessedData/DOC_TOTN_Asa_2024_2024.csv")
DOC_TOTN_Asa_2025_2026 <- read_csv("ProcessedData/DOC_TOTN_Asa_2025_2026.csv")

#Nitrogen and Phosphorous NP
NP_Asa_2022_2022 <- read_csv("ProcessedData/NP_Asa_2022_2022.csv")
NP_Asa_2023_2023 <- read_csv("ProcessedData/NP_Asa_2023_2023.csv")
NP_Asa_2023_2024 <- read_csv("ProcessedData/NP_Asa_2023_2024.csv")
NP_Asa_2024_2025 <- read_csv("ProcessedData/NP_Asa_2024_2025.csv")
NP_Asa_2025_2025 <- read_csv("ProcessedData/NP_Asa_2025_2025.csv")

#EC and pH (sometimes combined with TagTag)
EC_pH_Asa_2022_2025 <- read_csv("ProcessedData/EC_pH_Asa_2022_2025.csv")


```


```{r}
# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "ASA" #STATION Abbreviation here
SamplingLocation <- "FER" #sampling code here
EndYear <- 2025 #END YEAR HERE
```


convert LOD to 99999 and NaN to NA and convert everything to numeric 
```{r}
TagTag <- TagTag %>%
  mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.numeric))

df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
df_names <- df_names[df_names != "TagTag"]

for(df_name in df_names) {
  assign(df_name, get(df_name) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.character)) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(. == "LOD", "99999", .))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ suppressWarnings(as.numeric(.)))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(is.nan(.), NA_real_, .)))
  )
}
```

merge any overlapping data sets
```{r}
#NP
NP_all <- bind_rows(mget(ls(pattern = "^NP"))) |> 
  distinct()
rm(list = setdiff(ls(pattern = "^NP"), "NP_all"))

#DOC
DOC_TOTN_all <- bind_rows(mget(ls(pattern = "^DOC")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^DOC"), "DOC_TOTN_all"))

#Anions
Anions_all <- bind_rows(mget(ls(pattern = "^Anions")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Anions"), "Anions_all"))

#Cations
Cations_all <- bind_rows(mget(ls(pattern = "^Cations")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Cations"), "Cations_all"))

```


merge data - keep everything from TagTag leave NAs in the other locations
```{r}
TagTag <- TagTag %>%
  left_join(DOC_TOTN_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(NP_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Anions_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Cations_all, by = "SampleID")

rm(DOC_TOTN_all, NP_all, Anions_all, Cations_all)

#just for abisko - drop F until things are re-run - there was a problem with a difference in magnitude between the two labs
# TagTag <- TagTag %>% 
#   select(-"F")

#just for Asa that doesn't have the pH and EC in the same file as the TagTag
TagTag <- TagTag %>%
  left_join(EC_pH_Asa_2022_2025, by = "SampleID")

rm(EC_pH_Asa_2022_2025)

```

#Data on Portal
##Upload data
for example Almbergasjön: SITES_CHEM-L_ANS_ALP_20170626-20220924_L2_irregular
SITES_CHEM-L_ASA_FER_20170622-20220905_L2_irregular.csv
```{r}
skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHEM-L_ASA_FER_20170622-20220905_L2_irregular.csv"), "####"))[1]
OnPortal <- read_csv("OnPortal/SITES_CHEM-L_ASA_FER_20170622-20220905_L2_irregular.csv", skip = skip_lines)

#change LOD to error code 99999 and NaN to NA and convert everything to numeric 
OnPortal <- OnPortal %>%
  mutate(across(!any_of("TIMESTAMP"), as.character)) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == "LOD", "99999", .))) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ suppressWarnings(as.numeric(.)))) %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(is.nan(.), NA_real_, .)))


#just for abisko 
# OnPortal <- OnPortal %>%
#   filter(TIMESTAMP <= as.Date("2022-04-05"))
```

**JUST ABISKO**
NaN from start until 2019 and 
LOD from 2019-09-17 onward
except for BR is 10.624 at 4m, 2020-04-03
```{r}
OnPortal <- OnPortal %>%
  mutate(
    Br = case_when(
      TIMESTAMP < as.Date("2019-09-04") ~ "NaN",
      TIMESTAMP >= as.Date("2019-09-17") ~ "LOD",
      TRUE ~ NA_character_))

OnPortal <- OnPortal %>%
  mutate(
    Br = ifelse(as.Date(TIMESTAMP) == as.Date("2020-04-03") & D_WATER_MIN == 4, 
                10.624, 
                Br))
```

**Just Asa**
```{r}
#drop the cations 2017-08-19 outliers across the board 
keep_prefixes <- c("EC", "pH", "TW", "DOC", "TOT-N", "PO4-P", "NH4-N", "NO3-N", "F", "Cl", "Br", "SO4-S")

# Identify columns to nullify (excluding TIMESTAMP)
nullify_cols <- names(OnPortal)[!grepl(paste0("^(", paste(keep_prefixes, collapse = "|"), ")"), names(OnPortal))]
nullify_cols <- nullify_cols[nullify_cols != "TIMESTAMP"]

# Set values to NA for the target timestamp
OnPortal <- OnPortal %>%
  mutate(across(all_of(nullify_cols), ~ if_else(TIMESTAMP == as.POSIXct("2017-08-09 12:00:00", tz = "UTC"), NA_real_, .)))

#drop the SO4-S that is currently on the Portal between 2020-10-12 12:00:00 and 2022-09-05 12:00:00
so4_cols <- names(OnPortal)[grepl("^SO4-S", names(OnPortal))]

OnPortal <- OnPortal %>%
  mutate(across(all_of(so4_cols), ~ if_else(
    TIMESTAMP >= as.POSIXct("2020-10-12 12:00:00", tz = "UTC") & 
    TIMESTAMP <= as.POSIXct("2022-09-05 12:00:00", tz = "UTC"), 
    NA_real_, .)))

#remove anions for 2018-04-23 12:00:00
anions <- names(OnPortal)[grepl("^(Cl|SO4-S)", names(OnPortal))]

OnPortal <- OnPortal %>%
  mutate(across(all_of(anions), ~ if_else(
    TIMESTAMP == as.POSIXct("2018-04-23 12:00:00", tz = "UTC"),
    NA_real_, .)))
```


##Reformat for preview
```{r}
#Only for Abisko
# OnPortal <- OnPortal %>% 
#   select(-D_WATER_MAX)

OnPortal_long <- OnPortal %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == 99999, NA_real_, .)))

#Pivot to long format - ABISKO
# OnPortal_long <- OnPortal_long |> 
#   pivot_longer(
#     cols = -c(TIMESTAMP, D_WATER_MIN), 
#     names_to = "Variable", 
#     values_to = "Value")

#Pivot to long format
OnPortal_long <- OnPortal_long |> 
  pivot_longer(
    cols = -c(TIMESTAMP), 
    names_to = "Variable", 
    values_to = "Value")

#for locations with depth in the header
OnPortal_long <- OnPortal_long %>%
  mutate(D_WATER = as.numeric(str_extract(Variable, "-[0-9.]+")),
         Variable = str_remove(Variable, "_-[0-9.]+m?"))
OnPortal_long <- OnPortal_long %>%
  relocate(D_WATER, .after = Variable)

#drop NAs
OnPortal_long <- OnPortal_long %>%
  drop_na(Value)

#add a column saying that this is the data from the portal
OnPortal_long <-  OnPortal_long %>% 
  mutate(dataset = "OnPortal")

```

reformat new data for preview
```{r}
#Abisko
# TagTag_long <- TagTag %>% 
#   select(-D_WATER_MAX)

#filter out just the lake data based on the SAMPLE ID 
#asa specific ending in 31 is 0.5m ending in 32 is 9m
# add a depth column
TagTag_long <- TagTag %>%
  filter(str_ends(SampleID, "31|32")) %>%
  mutate(D_WATER = case_when(
    str_ends(SampleID, "31") ~ -0.5,
    str_ends(SampleID, "32") ~ -9,
    TRUE ~ NA_real_
  ))

TagTag_long <- TagTag_long %>% 
  select(-SampleID)

TagTag_long <- TagTag_long %>%
  mutate(TIMESTAMP = as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"))

TagTag_long <- TagTag_long %>%
  mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == 99999, NA_real_, .)))

#Pivot to long format - ABISKO
# TagTag_long <- TagTag_long |> 
#   pivot_longer(
#     cols = -c(TIMESTAMP, D_WATER_MIN), 
#     names_to = "Variable", 
#     values_to = "Value")

#Pivot to long format
TagTag_long <- TagTag_long |> 
  pivot_longer(
    cols = -c(TIMESTAMP, D_WATER), 
    names_to = "Variable", 
    values_to = "Value")

TagTag_long <- TagTag_long %>%
  drop_na(Value)

#add a column saying that this is the new data
TagTag_long <-  TagTag_long %>% 
  mutate(dataset = "NewData")

TagTag_long <- TagTag_long %>%
  relocate(D_WATER, .after = Variable)
```


#Preview all data
```{r}
#combine the long formats of the old and new data here
chem_combined_long <- bind_rows(OnPortal_long, TagTag_long)

library(ggforce) #for the qc plots

pdf(paste0("QC_plots/", StationName, "_", SamplingLocation, "_QC_plots.pdf"), width = 11, height = 8.5)

n_pages <- nlevels(factor(chem_combined_long$Variable)) %/% 6 + 1 #if it divides evenly you don't need the plus one and it will throw and error but the error only means there is an extra page, which is fine. 

#in the model for Abisko - geom_line(aes(group = interaction(dataset, D_WATER_MIN)), linewidth = 0.3) +

for (i in seq_len(n_pages)) {

p <- ggplot(chem_combined_long,
    aes(TIMESTAMP, Value, 
        color = factor(D_WATER), 
        shape = dataset)) +
    geom_line(aes(group = interaction(dataset, D_WATER)), linewidth = 0.3) +
    geom_point(size = 1) +
    facet_wrap_paginate(
      ~ Variable,
      scales = "free_y",
      ncol = 2,
      nrow = 3,
      page = i) +
    scale_color_manual(
      values = c(
        "-0.5" = "grey60",
        "-9" = "black"),
      name = "Depth (m)") +
    scale_shape_manual(
      values = c(
        "OnPortal" = 4,   # x
        "NewData" = 16),  # closed dot
      name = "Dataset") +
    labs(
      title = paste("QC plots — page", i),
      x = "Date",
      y = "Value") +
    theme_light()
  print(p)   # one page per iteration
}

dev.off()

```

#Merge the old and new data for export
##Abisko and Erken which have max and min depth
```{r}
variable_order <- c("TIMESTAMP", "D_WATER_MIN", "D_WATER_MAX", "pH","DOC","TOT-N","PO4-P","NH4-N","NO3-N","Cl","SO4-S","Br","Li","Be","B","Na","Mg","Al","Si","P","S","K","Ca","V","Cr","Mn","Fe","Co","Ni","Cu","Zn","Ga","As","Se","Rb","Sr","Ag","Cd","In","Te","Cs","Ba","Tl","Pb","Bi")

TagTag <- TagTag %>% 
  select(-SampleID) %>%
  mutate(`SO4-S` = as.character(round(as.numeric(`SO4-S`), 4))) %>%
  select(all_of(variable_order)) %>%
  mutate(across(-TIMESTAMP, ~replace(., is.na(.), NaN)))

OnPortal <- OnPortal %>%
  mutate(across(-1, as.character))%>%
  select(all_of(variable_order))

LakeChemistry <- bind_rows(OnPortal, TagTag)

LakeChemistry <- LakeChemistry %>% 
mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))
```

#Asa
```{r}
LakeChemistry <- chem_combined_long |> 
  select(-dataset)

LakeChemistry <- LakeChemistry |> 
  pivot_wider(
    names_from = c(Variable, D_WATER),
    values_from = Value,
    names_glue = "{Variable}_{D_WATER}m"
  )

#this is what is in the file that is currently on the portal 
variable_order <- c("TIMESTAMP","EC_-0.5m","EC_-9m","pH_-0.5m","pH_-9m","TW_pH_-0.5m","TW_pH_-9m","DOC_-0.5m","DOC_-9m","TOT-N_-0.5m","TOT-N_-9m","PO4-P_-0.5m","PO4-P_-9m","NH4-N_-0.5m","NH4-N_-9m","NO3-N_-0.5m","NO3-N_-9m","F_-0.5m","F_-9m","Cl_-0.5m","Cl_-9m","Br_-0.5m","Br_-9m","SO4-S_-0.5m","SO4-S_-9m","Li_-0.5m","Li_-9m","Be_-0.5m","Be_-9m","B_-0.5m","B_-9m","Na_-0.5m","Na_-9m","Mg_-0.5m","Mg_-9m","Al_-0.5m","Al_-9m","Si_-0.5m","Si_-9m","P_-0.5m","P_-9m","S_-0.5m","S_-9m","K_-0.5m","K_-9m","Ca_-0.5m","Ca_-9m","V_-0.5m","V_-9m","Cr_-0.5m","Cr_-9m","Mn_-0.5m","Mn_-9m","Fe_-0.5m","Fe_-9m","Co_-0.5m","Co_-9m","Ni_-0.5m","Ni_-9m","Cu_-0.5m","Cu_-9m","Zn_-0.5m","Zn_-9m","Ga_-0.5m","Ga_-9m","As_-0.5m","As_-9m","Se_-0.5m","Se_-9m","Rb_-0.5m","Rb_-9m","Sr_-0.5m","Sr_-9m","Ag_-0.5m","Ag_-9m","Cd_-0.5m","Cd_-9m","Te_-0.5m","Te_-9m","Cs_-0.5m","Cs_-9m","Ba_-0.5m","Ba_-9m","Tl_-0.5m","Tl_-9m","Pb_-0.5m","Pb_-9m","Bi_-0.5m","Bi_-9m")

LakeChemistry <- LakeChemistry |>  
  select(all_of(variable_order)) |> 
  mutate(across(-TIMESTAMP, ~ ifelse(. == 99999, "LOD", .)),
         across(-TIMESTAMP, ~ replace(., is.na(.), NaN)))

LakeChemistry <- LakeChemistry %>% 
mutate(TIMESTAMP = format(as.POSIXct(TIMESTAMP, format = "%Y-%m-%d %H:%M"), "%Y-%m-%d %H:%M"))
```


```{r}
# Construct the filename
filename <- paste0("FinalChemFiles/CHEM-L_", StationName, "_", SamplingLocation, "_","until" ,EndYear, ".csv")

# Export to CSV
write.csv(LakeChemistry, filename, row.names = FALSE, quote = FALSE)
```

