---
title: "NP"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

load packages
```{r}
library(tidyverse)
library(readxl)
```

#SLU Umeå
NOTE - check the detection limits in the file to make sure they are the same as the header template
Create a new sheet called RESULTS - just copy paste the full table into it (starting with the green boxes)

load data - sometimes the filenames are different so you cant fully automate
*CHANGE SOMETHING MANUALLY HERE*
```{r}
rawfile <- "NP_SLU_Umeå_Asa_20241113" #ENTER YOUR FILENAME HERE

filepath <- paste0("LabResults/NP/", rawfile, ".xlsx")
df <- read_excel(filepath, sheet = "RESULTS")

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "Asa" #STATION NAME HERE
StartYear <- 2023 #START YEAR HERE
EndYear <- 2024 #END YEAR HERE
```

*check results*
if something other than ug/L shows up you will need to do a conversion
```{r}
unique(as.character(unlist(df[1, ])))
```

as long as the above is µg/L - continue
```{r}
#drop the first row
df <- df %>%
  slice(-1)

#drop unnecessary columns
df <- df %>%
  select(-any_of(c("CUSTID 1", "CUSTID 2", "KUNDID 1", "KUNDID")))
```

*check data*
Check whether there are any that = 9999 (using this as the error code)
```{r}
df %>%
  summarise(across(everything(), ~ any(. == 9999, na.rm = TRUE)))
```


*check results*
Check whether there are any LODs in the results - if there are you need to convert them to 9999 and then convert them back later 
```{r}
df %>%
  summarise(across(everything(), ~ any(grepl("^<$|^LOD$|^<LOD$", ., ignore.case = TRUE), na.rm = TRUE))) #this shows anything that has < or LOD or <LOD, this should also catch any weird spaces in the LOD cell

#only need to run this if there are LOD above
df <- df %>%
  mutate(across(everything(), ~ ifelse(grepl("^<|^LOD$", ., ignore.case = TRUE), 9999, .)))
```

convert to numeric
```{r}
df <- df %>%  # convert all columns except the first to numeric
 mutate(across(-1, as.numeric))
```

*check results*
This is looking for values below LOD -- that need to be converted to LOD -- if there are any you need to add more code
You want it to say numeric(0) - then you don't need to do anything else
```{r}
df %>%
  filter(`P-PO4` < 0.4) %>%
  pull(`P-PO4`)

df %>%
  filter(`N-NH4` < 0.3) %>%
  pull(`N-NH4`)

df %>%
  filter(`N-NO3` < 0.4) %>%
  pull(`N-NO3`)
```

change names
```{r}
df <- df %>%
  rename(
    SampleID = `LABID`,
    `PO4-P` = `P-PO4`,
    `NH4-N` = `N-NH4`,
    `NO3-N` = `N-NO3`)
```

*convert anything with 9999 to LOD
```{r}
df <- df %>%
  mutate(across(-SampleID, ~ ifelse(. == 9999, "LOD", .)))
```


Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/NP_", StationName, "_", StartYear,"_",EndYear, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE, quote = FALSE)
```
