---
title: "Skogaryd_Chemistry_Merge"
author: "Matilda Andersson"
date: "2026-06-24"
output: html_document
---

NOTE SRS is a bit annoying because they have multiple datasets for past years due to shifts in the method and/or lab


load packages
```{r}
library(tidyverse)
library(readxl)
```

load data - this now just uploads the entire Abisko folder with all the data
IMPORTANT - it's important that the file names follow the naming conventions so that everything uploaded makes sense
```{r}
list.files(path = "ProcessedData/Skogaryd", 
           pattern = "\\.csv$", 
           full.names = TRUE) %>%
  set_names(~ basename(.) %>% str_remove("\\.csv$")) %>%
  map(read_csv) %>%
  list2env(envir = .GlobalEnv)

#you might need to merge multiple TagTag files and rename
TagTag <- TagTag_Skogaryd_2024

rm(TagTag_Skogaryd_2024)
```

```{r}
# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "SRC" #STATION Abbreviation here
LakeSamplingLocation <- "ERS" #sampling code here
EndYear <- 2025 #END YEAR HERE
```

convert LOD to 99999 and NaN to NA and convert everything to numeric 
```{r}
TagTag <- TagTag %>%
  mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.numeric))

#this excluded TagTag
df_names <- ls()[sapply(ls(), function(x) is.data.frame(get(x)))]
df_names <- df_names[df_names != "TagTag"]

for(df_name in df_names) {
  assign(df_name, get(df_name) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), as.character)) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(. == "LOD", "99999", .))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ suppressWarnings(as.numeric(.)))) %>%
    mutate(across(!any_of(c("TIMESTAMP", "SampleID")), ~ if_else(is.nan(.), NA_real_, .)))
  )
}
```

merge any overlapping data sets
```{r}
#NP
NP_all <- bind_rows(mget(ls(pattern = "^NP"))) |> 
  distinct()
rm(list = setdiff(ls(pattern = "^NP"), "NP_all"))

#DOC
DOC_TOTN_all <- bind_rows(mget(ls(pattern = "^DOC")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^DOC"), "DOC_TOTN_all"))

#Anions
Anions_all <- bind_rows(mget(ls(pattern = "^Anions")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Anions"), "Anions_all"))

#Cations
Cations_all <- bind_rows(mget(ls(pattern = "^Cations")))|> 
  distinct()
rm(list = setdiff(ls(pattern = "^Cations"), "Cations_all"))

```

merge data - keep everything from TagTag leave NAs in the other locations
```{r}
TagTag <- TagTag %>%
  left_join(DOC_TOTN_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(NP_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Anions_all, by = "SampleID")

TagTag <- TagTag %>%
  left_join(Cations_all, by = "SampleID")

rm(DOC_TOTN_all, NP_all, Anions_all, Cations_all)
```

#Lake
##Data on Portal
Upload data - you need to download the most recent file on the portal for this
```{r}
skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHE-L_SRC_ERS_20160510-20170328_L2_irregular.csv"), "####"))[1]
OnPortal_till2017 <- read_csv("OnPortal/SITES_CHE-L_SRC_ERS_20160510-20170328_L2_irregular.csv", skip = skip_lines)

skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHE-L_SRC_ERS_20170515-20191112_L2_irregular.csv"), "####"))[1]
OnPortal_till2019 <- read_csv("OnPortal/SITES_CHE-L_SRC_ERS_20170515-20191112_L2_irregular.csv", skip = skip_lines)

skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHE-L_SRC_ERS_20191218-20200611_L2_irregular.csv"), "####"))[1]
OnPortal_till2020 <- read_csv("OnPortal/SITES_CHE-L_SRC_ERS_20191218-20200611_L2_irregular.csv", skip = skip_lines)

skip_lines <- which(str_detect(readLines("OnPortal/SITES_CHE-L_SRC_ERS_20200622-20221019_L2_irregular.csv"), "####"))[1]
OnPortal_till2022 <- read_csv("OnPortal/SITES_CHE-L_SRC_ERS_20200622-20221019_L2_irregular.csv", skip = skip_lines)

#change LOD to error code 99999 and NaN to NA and convert everything to numeric -- you need to do this before rbind because it wont join the columns if they aren't the same type
#unlike for the other stations that only have 1 file for lake chemistry, this is a loop so that it can be applied to all the dfs
clean_df <- function(df) {
  df %>%
    mutate(across(!any_of("TIMESTAMP"), as.character)) %>%
    mutate(across(!any_of("TIMESTAMP"), ~ if_else(. == "LOD", "99999", .))) %>%
    mutate(across(!any_of("TIMESTAMP"), ~ suppressWarnings(as.numeric(.)))) %>%
    mutate(across(!any_of("TIMESTAMP"), ~ if_else(is.nan(.), NA_real_, .)))
}

# Apply to all four and overwrite originals
list("OnPortal_till2017", "OnPortal_till2019", "OnPortal_till2020", "OnPortal_till2022") %>%
  walk(~ assign(.x, clean_df(get(.x)), envir = .GlobalEnv))

#Merge all the old chemistry data
OnPortal <- bind_rows(OnPortal_till2017, OnPortal_till2019, OnPortal_till2020, OnPortal_till2022)

rm(OnPortal_till2017, OnPortal_till2019, OnPortal_till2020, OnPortal_till2022)
```

```{r}

```

