---
title: "TagTag"
author: "Matilda Andersson"
date: "2025-10-01"
output: html_document
---

NOTE
ABISKO CODE AT THE BOTTOM
For the older abisko data there isn't TagTag so there is specific code at the bottom to deal with that

Load packages
```{r}
library(tidyverse)
library(readxl)
library(lubridate)
```

load data - sometimes the filenames are different so you cant fully automate
*CHANGE SOMETHING MANUALLY HERE*
```{r}
rawfile <- "5-101F TagTags export_Skogaryd_2024" #ENTER YOUR FILENAME HERE

filepath <- paste0("TagTag/SRC/", rawfile, ".xlsx")
df <- read_excel(filepath)

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "Skogaryd" #STATION NAME HERE
Year <- 2024 #START YEAR HERE
```

select columns to keep
```{r}
df <- df |> 
  select(Identifier,date, time, hach_ph, hach_ph_temp, hach_conductivity, hach_conductivity_temp, o2_mgl, o2_sat, o2_temp)
```

```{r}
df <- df %>%
  mutate(across(-(1:3), as.numeric))
```

```{r}
#drop any row with an NA in date
df <- df %>%
  filter(!is.na(date))

#fill empty cells in time
df <- df %>%
  mutate(time = replace_na(time, as.POSIXct("1899-12-31 12:00:00")))

df <- df %>%
  mutate(TIMESTAMP = ymd_hms(paste(as.Date(date), 
                                   format(time, "%H:%M:%S"))))

df <- df |> 
  select(-date,-time)

df <- df %>%
  rename(
    SampleID = `Identifier`,
    pH = hach_ph,
    TW_pH = hach_ph_temp, 
    EC = hach_conductivity, 
    TW_EC = hach_conductivity_temp, 
    O2 = o2_mgl,
    O2SAT = o2_sat,
    TW_O2 = o2_temp)

df <- df %>%
  relocate(TIMESTAMP, .after = SampleID)
```

Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/", StationName, "/TagTag_", StationName, "_", Year, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE)
```

```{r}
rm(list = ls())
```


#Abisko - before TagTag
*CHANGE SOMETHING MANUALLY HERE*
```{r}
rawfile <- "Almberga_grab_sample_LOG_from2022" #ENTER YOUR FILENAME HERE

filepath <- paste0("TagTag/2024/", rawfile, ".xlsx")
df <- read_excel(filepath)

# WRITE YOUR STATION NAME, START YEAR AND END YEAR ACCORDING TO WHAT IS IN YOUR DATA
StationName <- "ANS" #STATION NAME HERE
SamplingLocation <- "ALM" #SAMPLING LOCATION HERE (if there are different logs for each location)
Year <- 2022 #START YEAR HERE
```

clean up and re-name
```{r}
#duplicate the depth column (for the early years they had integrated samples)
df <- df %>%
  mutate(
    D_WATER_MIN = `Sampling depth`,
    D_WATER_MAX = `Sampling depth`) %>%
  select(-`Sampling depth`)

df <- df %>%
  rename(
    SampleID = `QR code`,
    pH = `pH (25C)`)

df <- df %>%
  relocate(TIMESTAMP, .after = SampleID)
```

Write clean data to .csv
```{r}
# Construct the filename
filename <- paste0("ProcessedData/TagTag_", StationName, "_", SamplingLocation, "_", Year, ".csv")

# Export to CSV
write.csv(df, filename, row.names = FALSE)
```



